---
title: "data_analysis_covid"
author: "ELharith Mohamed"
date: "2022-10-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Cleaning your workspace
```{r}
rm(list = ls())
```

# Load my libraries

```{r}
library(tidyverse)
library(reshape2)
library(lubridate)
library(readxl)

```

# Loading covid-19 data
```{r}
tb <- read_csv("time_series_covid19_confirmed_global.csv")

```

# Cleaning Our Dataset
```{r}
# Dropped the Uneeded columns
tb_1 <- subset(tb, select = -c (`Province/State`, Lat, Long)) 

# Choosing our desired Countries data 
tb_2 <- filter(tb_1, tb_1$`Country/Region` == "Egypt" | tb_1$`Country/Region` == "Qatar")

# Transposing our data & adding it to a data frame
tb_3 <- as.data.frame(t(tb_2[,2:length(tb_2[1,])])) 

# naming our columns in the new data frame
names(tb_3) <- t(tb_2[,1])

# Exracting the dates from our Data frame
current_dates <- names(tb_2)[2:length(tb_2[1,])]

# adding the date column to my data frame
tb_3$date <- mdy(current_dates)

tail(tb_3)

```



```{r}
# Finding the daily counts from the cumlative data in Egypt

egypt_daily_counts <- c(0, diff(tb_3[,1]))
qatar_daily_counts <- c(0, diff(tb_3[,2]))

# Creating a data frame for Egypt's Data
egypt_daily_cases <- data.frame(Daily_cases = egypt_daily_counts, date = tb_3$date)

# Creating a data frame for Qatar's Data
qatar_daily_cases <- data.frame(Daily_cases = qatar_daily_counts, date = tb_3$date)


tail(egypt_daily_cases)

```

# Checking Our Data 
```{r}

nrow(egypt_daily_cases)
nrow(qatar_daily_cases)

```

```{r}
x_1 = egypt_daily_cases[(duplicated(egypt_daily_cases$date)),]
x_2 = qatar_daily_cases[(duplicated(qatar_daily_cases$date)),]
x_1
x_2

```

# Plotting Our Covid-19 data in Egypt & Qatar
```{r}
## Plotting Egypt's Data
egyptplot <- ggplot(egypt_daily_cases, aes(x=date, y = Daily_cases))

egyptplot + geom_point() + stat_smooth(span = 0.3) + ggtitle(paste("Daily Cases in Egypt"))

## Plotting Qatar's Data
qatarplot <- ggplot(egypt_daily_cases, aes(x=date, y = Daily_cases))

qatarplot + geom_point() + stat_smooth(span = 0.3) + ggtitle(paste("Daily Cases in Qatar"))

```
# Loading Google Mobility data in Egypt
```{r}

eg_mob_2020 <- read_csv("Region_Mobility_Report_CSVs/2020_EG_Region_Mobility_Report.csv")
eg_mob_2021 <- read_csv("Region_Mobility_Report_CSVs/2021_EG_Region_Mobility_Report.csv")
eg_mob_2022 <- read_csv("Region_Mobility_Report_CSVs/2022_EG_Region_Mobility_Report.csv")
eg_mob_all <- rbind(eg_mob_2020, eg_mob_2021,eg_mob_2022)
tail(eg_mob_all)
nrow(eg_mob_all)

```





```{r}
# Loading Google Mobility data in Qatar

qa_mob_2020 <- read_csv("Region_Mobility_Report_CSVs/2020_QA_Region_Mobility_Report.csv")
qa_mob_2021 <- read_csv("Region_Mobility_Report_CSVs/2021_QA_Region_Mobility_Report.csv")
qa_mob_2022 <- read_csv("Region_Mobility_Report_CSVs/2022_QA_Region_Mobility_Report.csv")
qa_mob_all <- rbind(qa_mob_2020,qa_mob_2021,qa_mob_2022)
tail(qa_mob_all)
nrow(qa_mob_all)

```

# Cleaning Our Data By Removing rows with country_region = NA
```{r}
eg_mob_all_2 <- eg_mob_all[is.na(eg_mob_all$country_region),]

qa_mob_all_2 <- qa_mob_all[is.na(qa_mob_all$country_region),]

eg_mob_all_2
qa_mob_all_2

```
# Checking Duplicated data on Egypt & Qatar Data
```{r}
e <- eg_mob_all[duplicated(eg_mob_all$date),]
e <- e[order(e$date, decreasing = FALSE),]

x <- qa_mob_all[duplicated(qa_mob_all$date),]
x <- x[order(x$date, decreasing = FALSE),]

e
qa_mob_all[(qa_mob_all$date == "2020-02-15	"),]

```
# Removing Duplicated data 
```{r}

gg_tbl <- eg_mob_all %>% group_by(date) %>% summarise_if(is.numeric, sum,na.rm =TRUE)

gg_tbll <- qa_mob_all %>% group_by(date) %>% summarise_if(is.numeric, sum,na.rm =TRUE)

nrow(gg_tbl)
nrow(gg_tbll)

```
# Merging Egypt Data 
```{r}
eg_merged_data <- merge(x = gg_tbl, y = egypt_daily_cases, by= "date", all.y = TRUE)
nrow(eg_merged_data)

```

# Merging Qatar Data 
```{r}
qa_merged_data <- merge(x = gg_tbll, y = qatar_daily_cases, by= "date", all.y = TRUE)
nrow(qa_merged_data)

```


# standardization step  for Egypt's Data 
```{r}
eg_merged_data_c2 <- eg_merged_data %>% dplyr::select(2:8)

# Scaling our Data / standardization step 

eg_merged_data_c2_scaled <- data.frame(scale(eg_merged_data_c2, center = TRUE, scale = TRUE))

eg_merged_data_c2_scaled$date = eg_merged_data$date

nrow(eg_merged_data_c2_scaled)
```

# standardization step  for Qatar's Data 
```{r}

qa_merged_data_c2 <- qa_merged_data %>% dplyr::select(2:8)

# Scaling our Data / standardization step 

qa_merged_data_c2_scaled <- data.frame(scale(qa_merged_data_c2, center = TRUE, scale = TRUE))

qa_merged_data_c2_scaled$date = qa_merged_data$date

head(qa_merged_data_c2_scaled)

```


# Selecting our Columns 
```{r}
select_columns <- c("Daily_cases","workplaces_percent_change_from_baseline",  "residential_percent_change_from_baseline")
```


# Plotting Egypt's Data 
````{r}
eg_melt <- melt(eg_merged_data_c2_scaled, measure.vars = select_columns, value.name = "Value", variable.name = "variable")
eg_melt_2 <- eg_melt[!(is.na(eg_melt$Value)),]
nrow(eg_melt_2)
eg_plot <- ggplot(eg_melt_2, aes(x = date, y = Value, color = variable)) + stat_smooth(span = 0.15)
eg_plot  + ggtitle(paste("Egypt Data"))

```

# Plotting Qatar's Data 

```{r}

qa_melt <- melt(qa_merged_data_c2_scaled, measure.vars = select_columns, value.name = "Value", variable.name = "variable")
qa_melt_2 <- qa_melt[!(is.na(eg_melt$Value)),]

qa_plot <- ggplot(qa_melt_2, aes(x = date, y = Value, color = variable)) + stat_smooth(span = 0.15)
qa_plot  + ggtitle(paste("Qatar Data"))

```



```{r}


```

